Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-09-30 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:09:30	15:06:29	(Wed30.Sep)

Adding slightly steered versions of each window to the window database gives LCA greater flexibility in searching for optimal window. A Kaiser window $\f_\beta$ steered to the angle $\phi$ can be expressed as 
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}
\end{align}
%
where diag($\a_\phi$) is a diagonal matrix constructed from the steering vector $\a_\phi$:
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{2\pi d}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{2\pi (M-1)d}{\lambda}\sin(\phi)}
}.\label{steering_vector}
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_definition}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

\subsection{Steering bounds}

@@

We call the upper bound for steering $\phi_{\text{3dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a -3dB crossing that is exactly half that of the unsteered window. Using a numerical search for the steering angle $\phi$ that fulfilled this criterion for a wide range of $\beta$-values, we obtained a third degree polynomial fit to these points:
%
\begin{align}
\phi_\text{3dB}(\beta) \approx \minus{}0.002894\beta^3 + 0.02451\beta^2 - 0.01308\beta + 0.2980
\end{align}
%
Henceforth any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

\subsection{Using trigonometric windows instead of Kaiser}

\begin{align}
\phi_\text{3dB}(\beta) \approx \minus{}2.214\beta^3 + 6.084\beta^2 - 5.710\beta + 2.136
\end{align}

@@

Throughout this article we have used LCA with variations of the Kaiser window function. However, we could instead have used a trigonometric window function:
%
\begin{align}
\w = e^{j\phi_\theta[m]}\big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\big)\label{trig_window_function}
\end{align}
%
where $\phi_\theta[m] = \frac{2\pi}{\lambda}dm\sin\theta$ is the steering angle. For this window there is an analytic solution for the optimal value of $\alpha$ in the minimum variance sense. To derive it, we start by inserting the window function (\ref{trig_window_function}) into the beamformer equation (\ref{z}):
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} e^{j\phi_\theta[m]}\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha \a\T\x + (1-\alpha)\b\T\x
\end{align}
%
where $\a$ is the steering vector defined in (\ref{steering_vector}), and 
%
\begin{align}
\b = \text{diag}(\a)\cdot\begin{bmatrix}
     1 &
     \cos\left(\frac{2\pi}{M-1}\right) &
     \cos\left(\frac{2\pi 2}{M-1}\right) &
     \hdots &
     1
     \end{bmatrix}\T.
\end{align}
%
To attain unity gain in the look direction the weights must sum to one:
%
\begin{align}
\w\T\1 &= \big(\alpha\a - (1-\alpha)\b\big)\T\1 = 1
\end{align}
%
This is true for any value of $\alpha$ if $\a\T\1 = 1$ and $\b\T\1 = 1$. Hence, we can preserve the unity gain constraint as long as we normalize $\a$ and $\b$.

>>>>

Adding slightly steered versions of each window to the window database gives LCA greater flexibility in searching for optimal window. A Kaiser window $\f_\beta$ steered to the angle $\phi$ can be expressed as 
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}\label{steered_window}
\end{align}
%
where diag($\a_\phi$) is a diagonal matrix constructed from the steering vector $\a_\phi$:
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{2\pi d}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{2\pi (M-1)d}{\lambda}\sin(\phi)}
}.\label{steering_vector}
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_definition}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

\subsection{Steering bounds}\label{steering_bounds}

@@

We call the upper bound for steering $\phi_{\text{3dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a -3dB crossing that is exactly half that of the unsteered window. Using a numerical search for the steering angle $\phi$ that fulfilled this criterion for a wide range of $\beta$-values, we obtained a third degree polynomial fit to these points:
%
\begin{align}
\phi_\text{3dB}(\beta) \approx \minus{}0.002894\beta^3 + 0.02451\beta^2 - 0.01308\beta + 0.2980\label{steering_kaiser_3dB}
\end{align}
%
Henceforth any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

\subsection{Trigonometric windows}

While we focus on using LCA with Kaiser windows, we will for completeness compare the performance with that of using trigonometric windows. This is defined as:
%
\begin{align}
f_m(\alpha) = \alpha - (1-\alpha)\cos\Big(\frac{2\pi m}{M-1}\Big)\label{trig_window_function}
\end{align}
%
where we apply the contraint $\alpha\in[0.5,1]$ to avoid windows with negative coefficients. Steering is applied as in (\ref{steered_window}), and the $\alpha$-value that halves the 3\,dB distance is found using the method described in section \ref{steering_bounds}:
%
\begin{align}
\phi_\text{3dB}(\alpha) \approx \minus{}2.214\beta^3 + 6.084\beta^2 - 5.710\beta + 2.136.
\end{align}

@@

Throughout this article we have used LCA with variations of the Kaiser window function. However, we could instead have used a trigonometric window function:

\ref{trig_window_function} \ref{steering_bounds}

%
\begin{align}
f_m(\alpha) = e^{j\phi_\theta[m]}\Big[\alpha - (1-\alpha)\cos\Big(\frac{2\pi m}{M-1}\Big)\Big]\label{trig_window_function}
\end{align}
%
where $\phi_\theta[m] = \frac{2\pi}{\lambda}dm\sin\theta$ is the steering angle. For this window there is an analytic solution for the optimal value of $\alpha$ in the minimum variance sense. To derive it, we start by inserting the window function (\ref{trig_window_function}) into the beamformer equation (\ref{z}):
%
\begin{align}
\w\H\x &= \sumb{m=0}{M-1} e^{j\phi_\theta[m]}\Big(\alpha - (1-\alpha)\cos\left(\frac{2\pi m}{M-1}\right)\Big) x_m^* \nn
&= \alpha \a\T\x + (1-\alpha)\b\T\x
\end{align}
%
where $\a$ is the steering vector defined in (\ref{steering_vector}), and 
%
\begin{align}
\b = \text{diag}(\a)\cdot\begin{bmatrix}
     1 &
     \cos\left(\frac{2\pi}{M-1}\right) &
     \cos\left(\frac{2\pi 2}{M-1}\right) &
     \hdots &
     1
     \end{bmatrix}\T.
\end{align}
%
To attain unity gain in the look direction the weights must sum to one:
%
\begin{align}
\w\T\1 &= \big(\alpha\a - (1-\alpha)\b\big)\T\1 = 1
\end{align}
%
This is true for any value of $\alpha$ if $\a\T\1 = 1$ and $\b\T\1 = 1$. Hence, we can preserve the unity gain constraint as long as we normalize $\a$ and $\b$.
